\section{Theory of Constrained Optimization}
The formulation of constrained problems in this book,
\begin{equation}
    \underset{x\in \mathbb{R}^n}{\min} f(x) \qquad \text{subject to} \quad 
    \begin{cases}
        c_i(x) &= 0, \quad i\in\mathcal{E},\\
        c_i(x) &\geq 0, \quad i\in\mathcal{I}
    \end{cases}
    \label{eq:constrained-problem}
\end{equation}
is standard in the literature, except for the sign chosen of the inequality constraint functions ($i\in\mathcal{I}$) which may vary from author to author. 
By defining the feasible set as
\begin{equation}
    \Omega \equiv \left\{ x \in \mathbb{R}^n | c_i(x) = 0, \forall i\in\mathcal{E}, c_i(x) \geq 0, \forall i \in \mathcal{I} \right\}
    \label{eq:feasible-set}
\end{equation}
the authors present a very elegant formulation of the constrained problem as
\[
    \underset{x\in\Omega}{\min}f(x).
\]
\subsection{Farkas' lemma}
In the proof of the first order optimality conditions (Theorem 12.1 in the book) the authors use
\begin{Lma}[Farkas]
    Let the cone K be defined as 
    \[
        K = \{By + Cw|y\geq 0\}.
    \]
    Given any vector $g\in\mathbb{R}^n$ we have either that $g\in K$ (i.e. $g = By + Cw$ for some $y\geq 0, w \in \mathbb{R}^n$) or that there exists $d\in\mathbb{R}^n$ satisfying 
    \begin{equation}
        g^T d < 0, \quad B^Td \geq 0, \quad C^T d = 0
        \label{eq:farkas_conditions}
\end{equation}
\end{Lma}
I felt the rest of the proof, after Farkas' lemma, was a bit thin so I thought I would expand a bit here.
The proof assumes the constraint qualification LICQ, that the gradients of the active constraints $\nabla c_i(x^*), \> i\in \mathcal{A}(x^*)$ at the optimum are linearly independent. 
Then they show that there exists a vector $\lambda^*$ of Lagrange multipliers which satisfies the KKT conditions.
They have previously showed that for all vectors $d$ in the tangent cone at the optimum $T_\Omega(x^*)$, the following holds
\[
    d^T\nabla f(x^*) \geq 0.
\]
They also show that if the LICQ hold, $T_\Omega(x^*) = \mathcal{F}(x^*)$. Recalling the definition of $\mathcal{F}(x^*)$, the set of first order feasible directions,
\[
    \mathcal{F}(x^*) = \left\{ d \> \left\vert \begin{aligned}
        \quad&d^T \nabla c_i(x^*) \geq 0,\quad \forall i \in A_\mathcal{I}(x^*)\cap\mathcal{I}\\
        \quad&d^T \nabla c_i(x^*) = 0, \quad \forall i\in\mathcal{E}
    \end{aligned}\right.\right\}
\]
By identifying $B$ with the inequality constraint gradients and $C$ with the equality constraint gradients in Farkas' lemma, we see that there must be some vector $\lambda_i^* \geq 0, i \in \mathcal{A}(x^*) \cap \mathcal{I}$ for which 
\[
    \nabla f(x^*) = \nabla c_i(x)\lambda^*
\]
since all vectors $d$ that satisfy the latter two of the conditions in \eqref{eq:farkas_conditions} do not satisfy $d^T\nabla f(x^*) < 0$.
The proof is then finished by verifying that this $\lambda^*$, together with the local solution $x^*$ does in fact satisfy all the KKT conditions.
\subsection{Another proof of the sufficient conditions}
I found the proof of the sufficient conditions to be somewhat confusing, so I decided to include a similar proof found in the book ''Foundations of Optimization'' by Osman Güler.
We begin by stating the 
\begin{Thm}[Second-order sufficient conditions]
Let $x^*$ be a feasible point for the problem \eqref{eq:constrained-problem} that satisfies the KKT conditions with Lagrange multiplier $\lambda^*$.
If 
\be
    d^T\nabla_x^2\mathcal{L}(x^*, \lambda^*) d > 0
    \label{eq:positive-definite-hessian-of-lagrangian}
\ee
for all $d\neq 0$ for which the following conditions hold\footnote{
    This set is also known as the critical cone at $x^*$, denoted $\mathcal{C}(x^*, \lambda^*)$.
    The critical cone is the set of directions $d\in\mathcal{F}(x^*)$ for which it is unknown if the objective will increase or decrease from first-order information only.
}
\begin{align*}% TODO Reverse this list.
    \nabla c_i(x^*)^T d =& 0, \quad \forall i\in\mathcal{E},\\
    \nabla c_i(x^*)^T d =& 0, \quad \forall i\in\mathcal{I}\cap\mathcal{A}(x^*) \text{ with } \lambda^*_i > 0\\
    \nabla c_i(x^*)^T d \geq& 0, \quad \forall i\in\mathcal{I}\cap\mathcal{A}(x^*),
\end{align*}
then $x^*$ is a strict local minimizer of \eqref{eq:constrained-problem} and there exists a constant $c$ and an open ball $B(x^*, \epsilon)$ such that
\be
    f(x) \geq f(x^*) + c\|x - x^*\|^2, \quad \forall x\in B(x^*, \epsilon) \cap \Omega
    \label{eq:sufficient-condition-statement}
\ee
\end{Thm}
\textit{Proof}. Suppose, for contradiction, there are some directions for which \eqref{eq:positive-definite-hessian-of-lagrangian} is satisfied, but not \eqref{eq:sufficient-condition-statement}.
Let $\epsilon_k$ be some positive sequence of numbers tending to 0 as $k$ goes to infinity, then there exists a sequence of feasible points $x_k \to x^*$, such that
\[ 
    f(x_k) < f(x^*) + \epsilon_k\|x^* - x_k\|^2.
\]
Define $d_k = x_k - x^*$ and let $d_k / \|d_k\| \to d$ with $\|d\| = 1$.
Then, on one hand we have
\begin{align*}
    f(x_k) &<\> f(x^*) + \epsilon_k \|d_k\|^2 + 0\\
    \epsilon_k\|d\|^2 &>\> f(x_k) - f(x^*) - \sum_{i\mathcal{A}(x^*)} \lambda^*_i c_i(x_k)\\
    &= \mathcal{L}(x_k, \lambda^*) - \mathcal{L}(x^*, \lambda^*)\\
    &= \nabla_x\mathcal{L}(x^*,\lambda^*)^T d_k + \frac{1}{2}d_k^T\nabla_x^2\mathcal{L}(x^*, \lambda^*)d_k + o(\|d_k\|^2)\\
    &= \frac{1}{2}d_k^T\nabla_x^2\mathcal{L}(x^*, \lambda^*)d_k + o(\|d_k\|^2)
\end{align*}
where first equality follows from $\sum \lambda_i^*c_i(x^*) = 0$ and the third from $\nabla_x\mathcal{L}(x^*,\lambda^*) = 0$, both of which are implied from the satisfaction of the KKT conditions.
Now, dividing both sides by $\|d_k\|^2 / 2$ and sending $k\to\infty$ gives the following 
\[
    d^T \nabla_x^2\mathcal{L}(x^*, \lambda^*) d \leq 0
\]
which violates \eqref{eq:positive-definite-hessian-of-lagrangian}.
On the other hand we have
\begin{align*}
    \epsilon_k \|d_k\|^2 &> f(x_k) - f(x^*) = \nabla f(x^*)^Td_k + o(\|d_k\|),\\
                       0 &\leq c_i(x_k) - c_i(x^*) = \nabla c_i(x^*)^T d_k + o(\|d_k\|), \quad i\in \mathcal{A}(x^*)\cap\mathcal{I}\\
                       0 &= c_i(x_k) - c_i(x^*) = \nabla c_i(x^*)^T d_k + o(\|d_k\|), \quad i \in \mathcal{E}
\end{align*}
Dividing each equation above by $\|d_k\|$ and sending $k\to\infty$ yields,
\[
    \nabla f(x^*)^Td \leq 0, \quad c_\mathcal{I}(x^*)^T d \geq 0, \quad c_\mathcal{E}(x^*)^T d = 0
\]
which along with $x^*$ satisfying the KKT conditions implies $d\in \mathcal{C}(x^*, \lambda^*)$.
We thus produced a vector $d$ in the critical cone violating \eqref{eq:positive-definite-hessian-of-lagrangian}, contradicting the assumptions of the theorem.\hfill\sq
% We assume that the hessian of the lagrangian is 'positive definite' for the directions in the critical cone
% Then we say that if that is the case - then we have a local solution.
% We assume the opposite, that we have a point for which it is true, but which is not a local solution.
% We then see that this leads to a contradiction, that for all directions d in the critical cone, that implies that 
% the hessian is 'negative definite' in that direction, leading to a contradiction. :) 
\subsection{Solutions}
The exercises chosen for this chapter are 12.1, 12.2, 12.4, 12.11 (not \textbf{b)}) and 12.22. 

\subsubsection{Exercise 12.1}
\begin{enumerate}
\item[(a)]
The feasible set is 
\[
    \Omega = \{x \in \mathbb{R} \mid c(x)=0\}
\]
The zeros of $c(x)$ are at $x = 0$ and at solutions of the equation $\sin\frac{1}{x} = 0$, which are $x_k = \frac{1}{k\pi}, \> k\in\mathbb{Z} \setminus \{0\}$.
To show $c(x) \in C^2$, we compute the first and second derivatives. For $x \neq 0$:
\begin{align*}
    c'(x) &= 6x^5 \sin(1/x) - x^4 \cos(1/x)\\
    c''(x) &= (30x^4-x^2)\sin(1/x) - 10x^3 \cos(1/x)
\end{align*}
As $x\to 0$, both $c'$ and $c''$ vanish, since $\sin, \cos$ remain bounded.
Thus, both first and second derivatives have well defined values in the entire domain, $c \in C^2$.

\item[(b)]
Let $x_k = 1/(k\pi)$ for $k \in \mathbb{Z}, k \neq 0$. The distance to the next feasible point $x_{k+1}$ is
\[
|x_k - x_{k+1}| = \left|\frac{1}{k\pi} - \frac{1}{(k+1)\pi}\right| = \frac{1}{\pi|k(k+1)|}
\]
Choose a neighborhood $\mathcal{N}_k$ of $x_k$ with radius $r = \frac{1}{2\pi|k(k+1)|}$. This neighborhood contains no other feasible points. Since $x_k$ is the only feasible point in $\mathcal{N}_k$, it is a local solution. As it is the only local solution in $\mathcal{N}_k$, it is an isolated local solution.

\item[(c)]
The objective function is $f(x) = x^2$. We have $f(0)=0$ and $f(x_k) = 1/(k\pi)^2 > 0$ for all $k \neq 0$. 
Since $f(0) < f(x_k)$ for all other feasible points $x_k$, $x=0$ is the global solution. 
It is also a strict local solution because for any neighborhood of $0$, there are feasible points $x_k$ for large enough $|k|$, and for these points $f(x_k) > f(0)$.
However, this also means that however small we choose our neighborhood around $0$, there will always be a $k$ large enough for other local solutions to exist within the neighborhood. 
Therefore, $x=0$ is not an isolated local solution.
\end{enumerate}
\subsubsection{Exercise 12.2}
An isolated solution is necessarily a strict local solution.
Recalling the
\begin{Def}[Isolated local solution]
    A point $x^*$ is called an \textit{isolated local solution} if $x^*\in\Omega$ and there is a neighborhood $\mathcal{N}$ of $x^*$ such that $x^*$ is the only local solution in $\mathcal{N}\cap\Omega$.
\end{Def}
This means that if a point $x^*$ is an isolated local solution, then there exists a neighborhood $\mathcal{N}$ where, for all points $x$, the following holds 
\[
    f(x^*) < f(x), \quad x, x^* \in \mathcal{N}
\]
which is precisely the condition for a strict local solution. 
It must be a strict inequality in the above condition since if equality held for some points in $\mathcal{N}$, they too would be local solutions, contradicting that $x^*$ was isolated.
A strict local solution is however not necessarily isolated as we saw in 12.1. There, $x = 0$ was the global optimum which means that for all $x$ in the feasible domain the following holds $f(0) < f(x), \quad x\in\Omega$, i.e. it is strict.
However, since there is no neighborhood where $x = 0$ is the only local solution, it cannot be isolated despite being strict.
\subsubsection{Exercise 12.4}
Just as in the proof of Theorem 2.5\footnote{Found on page 16 of the book.} we prove by contradiction.
Assume $x$ is a local but not global solution. Then there must exist some $z$ for which $f(z) < f(x)$. 
From the convexity of the feasible region, we can produce any convex combination of $x$ and $z$ and rest assured of its feasibility. 
So, if we consider the line between $x$ and $z$, which must be contained in $\Omega$ by convexity, we have from the convexity of $f$
\[
    f(\lambda z + (1 - \lambda)x) \leq \lambda f(z) + (1 - \lambda) f(x) < f(x)
\]
Thus, in an arbitrarily tiny neighborhood (radius $\lambda$) around $x$, there must exist points with a lower function value.
This contradicts the assumption that $x$ is a local solution and it must thus hold that any solution which is a local solution must also be a global.\hfill\sq 

\subsubsection{Exercise 12.11}
\begin{enumerate}
\item[(a)] The feasible set is $\Omega = \{ x \in \mathbb{R}^2 \mid x_2 \ge 0, x_2 \le x_1^2 \}$.
Let's define the constraint functions as $c_1(x) = x_2 \ge 0$ and $c_2(x) = x_1^2 - x_2 \ge 0$.
The tangent cone $T_\Omega(x^*)$ is the set of all directions $d$ for which there exists a sequence of feasible points $\{z_k\} \subset \Omega$, $z_k \to x^*$ and a sequence of positive scalars $\{t_k\}$, $t_k \to 0^+$ as $k\to\infty$ such that:
\[
d = \lim_{k \to \infty} \frac{z_k - x^*}{t_k}
\]
Now, let $d = (d_1, d_2)$ be a vector in $T_\Omega(x^*)$, and let $z_k = (x_{1k}, x_{2k})$. We can then write the above as:
\[
    (d_1, d_2) = \lim_{k \to \infty} \left( \frac{x_{1k}}{t_k}, \frac{x_{2k}}{t_k} \right)
\]
Since each $z_k \in \Omega$, it must satisfy the constraints. 
We use the constraints to furnish restrictions on the possible values of $d_1$ and $d_2$.
Starting out with $c_1$, we divide it by $t_k > 0$ to get the following condition:
\[
    \frac{x_{2k}}{t_k} \ge 0 \overset{k\to\infty}{\implies} d_2 \ge 0
\]
Dividing $c_2$ by $t_k$ as well, we get
\[
    \frac{x_{2k}}{t_k} \le \frac{x_{1k}^2}{t_k}.
\]
where can rewrite the right-hand side as:
\[
    \frac{x_{2k}}{t_k} \le \left( \frac{x_{1k}}{t_k} \right) \cdot x_{1k}.
\]
Now, take the limit as $k \to \infty$. We know $\lim \frac{x_{2k}}{t_k} = d_2$, $\lim \frac{x_{1k}}{t_k} = d_1$, and $\lim x_{1k} = 0$ (since $z_k \to x^*$).
\[
    d_2 \le \left( \lim_{k \to \infty} \frac{x_{1k}}{t_k} \right) \cdot \left( \lim_{k \to \infty} x_{1k} \right)
\]
\[
    d_2 \le d_1 \cdot 0
\]
\[
    d_2 \le 0
\]
Combining $d_2 \ge 0$ and $d_2 \le 0$, we must have $d_2 = 0$. There is no restriction on $d_1$.
Thus, finally, the tangent cone is
\[
    T_{\Omega}(x^*) = \left\{(d_1, 0) | d_1 \in \mathbb{R} \right\}.
\]
In the above, we have assumed that the tangent cone is not empty, in part \textbf{d)}, we show that is the case.

How about the set of first-order feasible directions, $\mathcal{F}$?
The gradients of the constraint functions,  $c_1(x) = x_2$, $c_2(x) = x_1^2 - x_2$, are
\begin{align*}
    \nabla c_1(x^*) &= (0, 1)\\
    \nabla c_2(x^*) &= (0, -1).
\end{align*}
The set of first-order feasible solutions is the set of directions $d$ (we only have inequality constraints in this case) for which the following conditions holds 
\[
    d^T \nabla c_i(x) \geq 0, \quad \forall i \in \mathcal{A}(x) \cap \mathcal{I}
\]
In order for $d^T\nabla c_i(x^*) \geq 0$ for both $i=0$ and $i=1$ we must have $d = (a, 0)^T$.
Thus, $\mathcal{F}(x^*) = T_\Omega(x^*)$.  
\item[(b)]
 *\textit{skip until next session}* LICQ are not satisfied since the gradients of the constraint functions are not linearly independent at $x^* = (0,0)^T$
\begin{align*}
    \nabla c_1(x^*) &= (0, 1)\\
    \nabla c_2(x^*) &= (0, -1)
\end{align*}
with $c_1(x) = x_2$, $c_2(x) = x_1^2 - x_2$.
\item[(c)]
If the objective function is $f(x) = -x_2$ the Lagrangian is
\begin{align*}
    \mathcal{L}(x, \lambda) = -x_2 - \sum_{i\in\{1,2\}} \lambda_i c_i(x) =& -x_2 -\lambda_1 x_2 - \lambda_2(x_1^2 - x_2)\\
    =& -\lambda_2x_1^2 + (\lambda_2 - \lambda_1 - 1)x_2
\end{align*}
So, (12.34a) becomes:
\[
    \nabla_x\mathcal{L}(x^*, \lambda^*) = \begin{pmatrix}
        0\\
        \lambda_2 - \lambda_1 - 1
    \end{pmatrix} \overset{!}{=} 0
\]
Choosing Lagrange multipliers such that $\lambda_2 - \lambda_1 = 1$ satisfies the above condition. 
Since the LICQ does not hold, $\lambda^*$ is not unique.
For (12.34b) and (12.34c) we already know $x^*$ is feasible.  
To satisfy (12.34d) we can pick $\lambda_1 = 1$, $\lambda_2 = 2$, for example.
Lastly, (12.34e) is satisfied since both constraint functions are zero at $x^*$.

\item[(d)]
\[
    z_k = \begin{pmatrix}
        t_k\\
        t_k^2
    \end{pmatrix}
\]
with $t_k = 1 / k$
\end{enumerate}
\subsubsection{Exercise 12.22}
We begin by making explicit the implications of the properties of $c_i(x), i\in\mathcal{E}$ being linear and $-c_i(x), i\in\mathcal{I}$ being convex. 
The defining property of a linear function is the following
\[
    f(ax + by) = af(x) + bf(y)
\]
and for a function to be convex means
\[
    f(\lambda x + (1-\lambda)y) \leq \lambda f(x) + (1 - \lambda) f(y),\qquad \lambda \in [0,1].
\]
In words, the function between two points $x,y$ is bounded above by the straight line connecting $f(x), f(y)$.
Additionally, the negative of a convex function is \textit{concave}, which has the same definition, but with $\leq$ switched for $\geq$.

Now for a set to be convex, we require any convex combination of points within the set to also be members of the set\footnote{A convex combination is a linear combination with non-negative coefficients summing to one. If you allow the coefficients to take any value, but still require them to sum to one, it is called an \textit{affine} combination.}.
That is,
\[
    \lambda x + (1 - \lambda)y = z \overset{!}{\in} \Omega
\]
To show this holds, we need to show $z$ is feasible.
\begin{align*}
    c_i(z) &= 0,\quad i\in\mathcal{E}\\
    c_j(z) &\geq 0, \quad i\in\mathcal{I}    
\end{align*}
Starting with the equality constraints
\begin{align*}
    0 \overset{!}{=} c_i(z) = c_i(\lambda x + (1-\lambda)y) = \lambda c_i(x) + (1-\lambda)c_i(y) = 0
\end{align*}
where the second inequality follows from linearity of $c_i, i\in\mathcal{E}$ and the third from $x,y\in\Omega$.
For the inequality constraints we have
\[
    0 \overset{!}{\leq} c_i(z) = c_i(\lambda x + (1-\lambda)y) \geq \lambda c_i(x) + (1-\lambda)c_i(y) \geq 0
\]
where the second inequality comes from $c_i, i \in \mathcal{I}$ being concave. \hfill\sq